---
title: 'SWEDEN: COVID-19 Analysis'
author: "David Jackson"
date: "02/15/2021 12:50 PM EST"
output: 
  html_document: 
    fig_width: 8
    fig_height: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,message = FALSE,warning = FALSE)
library(ggplot2)
library(scales)
library(RSQLite)
library(plotly)
library(dplyr)
library(forecast)
# theme_set(theme_linedraw() + theme(panel.grid = element_line(linetype = 'dashed')))
theme_set(theme_light())

```
```{r}
rm(list=ls())
source("../DATA/movavg.R")
db <- db <- dbConnect(RSQLite::SQLite(),dbname= "../COVID-19-DB/OURWORLD.sqlite3")
EU <- dbGetQuery(db,"select * from JHU ")
df <- dbGetQuery(db,"select * from JHU ")

df <- subset(df,location =="Sweden"  )
df$date <- as.Date(df$date)
EU$date <- as.Date(EU$date)
df <- df[order(df$date),]
EU <- EU[order(EU$date),]
df$CMA <- ma(df$new_cases,6,centre=TRUE)
df$DMA <- ma(df$new_deaths,6,centre=TRUE)
# dbDisconnect(db)
```

#### Example of : Daily data is compiled Tuesdays to Fridays 
* https://experience.arcgis.com/experience/09f821667ce64bf7be6f9f87457ed9aa/page/page_0/

```{r}
p1 <- df %>% filter(date >="2020-01-01" ) %>%
  ggplot() + geom_col(aes(x=date,y=new_cases,col="Daily Cases")) +
  labs(title="Cases: Plot of Tuesday - Friday Data.")

p2 <- df %>% filter(date >="2020-01-01" ) %>%
  ggplot() + geom_col(aes(x=date,y=CMA)) +
  labs(title="Cases : 7 Day Moving Average")

p3 <- df %>% filter(date >="2020-01-01" ) %>%
  ggplot() + geom_col(aes(x=date,y=new_deaths)) +
  labs(title="Deaths: Plot of Tuesday - Friday Data. ") +ylim(0,150)

p4 <- df %>% filter(date >="2020-01-01") %>%
  ggplot() + geom_col(aes(x=date,y=DMA)) +
  labs(title="Deaths: 7 Day Moving Average")
ggplotly(p1)
ggplotly(p2)
ggplotly(p3)
ggplotly(p4)
```

#### Daily COVID-19 Daily Cases

```{r}
p5 <- df %>% filter(date >="2020-02-27" & new_cases >0) %>%
ggplot() + geom_line(aes(x=date,y=new_cases,col="Daily Cases")) +
  geom_line(aes(x=date,y=CMA,col="7 Day Mov. Avg.")) +
  labs(title="Sweden Daily Daily cases")

p6 <- ggplot(df) + geom_col(aes(x=date,y=CMA,col=CMA)) + 
  labs(title="Seven Day Moving Average: Cases")
ggplotly(p5)
ggplotly(p6)
```

#### Daily Deaths

```{r}
p7 <-df %>% filter(date>="2020-03-20" & new_deaths>0) %>%
ggplot() +  geom_line(aes(x=date,y=new_deaths,col="Daily Deaths")) +
  geom_line(aes(x=date,y=DMA,col="7 Day Mov. Avg.")) +
  labs(title="Sweden Daily Deaths") 

p8 <- ggplot(df) + geom_col(aes(x=date,y=DMA,col=DMA)) + 
  labs(title="Seven Day Moving Average: Deaths")
ggplotly(p7)
ggplotly(p8)
  
```

#### Non-Moving Average By Week and By Month

```{r}
df$Monthly <- as.Date(cut(df$date,
  breaks = "month"))
df$Weekly <- as.Date(cut(df$date,
  breaks = "week",
  start.on.monday = FALSE))

```
```{r}
Weekly_new_cases <- aggregate(new_cases~Weekly,df,FUN=sum)
Weekly_new_deaths <- aggregate(new_deaths~Weekly,df,FUN=sum)
Weekly_new_cases$DRate <- Weekly_new_deaths$new_deaths/Weekly_new_cases$new_cases

```
```{r}
ggplot(Weekly_new_cases) + geom_col(aes(x=Weekly,y=new_cases)) + 
  labs(title="Weekly Cases",x="date date", y="Weekly Cases") +
  geom_hline(yintercept = mean(Weekly_new_cases$new_cases),col="red",lwd=1.5) +
  scale_y_continuous(labels = scales::comma) 


ggplot(Weekly_new_deaths) + geom_col(aes(x=Weekly,y=new_deaths)) + 
  labs(title="Weekly new_deaths",x="date date", y="Weekly Deaths") +
  scale_y_continuous(labels = scales::comma) +
  geom_hline(yintercept = mean(Weekly_new_deaths$new_deaths),col="red",lwd=1.5)
``` 

#### Monthly new_cases and new_deaths

```{r}

Monthly_new_cases <- aggregate(new_cases~Monthly,df,FUN=sum)
Monthly_new_deaths <- aggregate(new_deaths~Monthly,df,FUN=sum)
Monthly_new_cases$DRate <- Monthly_new_deaths$new_deaths/Monthly_new_cases$new_cases

```
```{r}
ggplot(Monthly_new_cases) + geom_col(aes(x=Monthly,y=new_cases)) +
  labs(title="Monthly Cases") +
  scale_y_continuous(labels = scales::comma)

ggplot(Monthly_new_deaths) + geom_col(aes(x=Monthly,y=new_deaths)) +
  labs(title="Monthly Deaths") +
  scale_y_continuous(labels = scales::comma)
```

#### Sweden in the EU: Covid-19 Cases and Deaths
* EU Counties with Population between 9,000,000 and 12,000,000

```{r}

EU1 <- subset(EU,location == c("Belgium","Czechia","Greece","Sweden","Hungary","Austria"))

EUTotal <-EU1 %>% group_by(location) %>% summarise(
                                Case =sum(new_cases,na.rm = TRUE),
                                Deaths = sum(new_deaths,na.rm = TRUE))
EUTotal$Rate <- EUTotal$Deaths/EUTotal$Case
```

#### Plots of Total Cases and Deaths

```{r}
ggplot(EUTotal) + geom_col(aes(x=reorder(location,Case),y=Case)) + 
  labs(title="Total Cases by Country",x="Countries",y="Total Cases") + scale_y_continuous(label=comma)

``` 
```{r}
ggplot(EUTotal) + geom_col(aes(x=reorder(location,Deaths),y=Deaths)) + 
  labs(title="Total Deaths by Country",x="Countries",y="Total  Deaths") + scale_y_continuous(label=comma)

```

#### Total Cases  by Day.

```{r}
EU2 <- EU1 %>% filter(date >="2020-09-01")
EU2 %>% ggplot(aes(x=date,y=new_cases,fill=location))+geom_area() + facet_wrap(~location,ncol = 2,scales="free_y") +
  expand_limits(y=0)
```

#### Total Deaths  by Day.

```{r}
EU2 <- EU1 %>% filter(date >="2020-09-01")
EU2 %>% ggplot(aes(x=date,y=new_deaths,fill=location))+geom_area() + facet_wrap(~location,ncol = 2,scales="free_y") +expand_limits(y=0)
  
```

### Cumlative Totals (Cases and Deaths) for Sweden
* Total Cases and Deaths to date

```{r}
df %>% ggplot() + geom_col(aes(x=date,y=total_cases,fill=new_cases)) +
  scale_y_continuous(labels = comma) +
  labs(title="Sweden Cumulative Cases by Date")

df %>% ggplot() + geom_col(aes(x=date,y=total_deaths,fill=new_deaths)) +
  scale_y_continuous(labels = comma) +
  labs(title="Sweden Cumulative Deaths by Date")
```

#### Plot by Country

```{r}
EU1 %>% ggplot() + geom_area(aes(x=date,y=total_cases,fill=location)) +
  scale_y_continuous(labels = comma) +
  labs(title=" Cumulative Cases by Country") +
  facet_wrap(~location,ncol = 2,scales="free_y")

EU1 %>% ggplot() + geom_area(aes(x=date,y=total_deaths,fill=location)) +
  scale_y_continuous(labels = comma) +
  labs(title="Cumulative Deaths by Country") +
  facet_wrap(~location,ncol = 2,scales="free_y")
```

